### 2023-11-03 Code/460ProjectDraft.Rmd
Working Draft

---
title: "460ProjectDraft"
author: "James Santarpio, Mardi"
date: "2023-11-03"
output: html_document
---
Project Description

Topic:
The effect of drug decriminilization on overdose deaths and well-being.

Why it matters:
The study of drug decriminalization and overdose rates in Oregon has potential for significant policy and societal relevance. There are many debates on the benefits and drawbacks of drugs use and decriminalization has potential for policy implications involving economics, public health, criminal justice, and community well-being.

Research Question and Motivation:
What is the effect of drug decriminalization on overdose deaths and well-being? A policy change like Oregon's can have significant consequences on populations who are most at risk of overdose deaths and the economic impacts on health care costs and overall well-being. We used the synthetic control method to understand how the policy change impacted Oregon by comparing what would have happened in th absence of the policy being implemented.

Description of Data:

Overdose Mortality by county in the US:
The Overdose Mortality by County in the US is from the CDC Center for Health Statistics for the years 2020-2023. These provisional county-Level drug Overdose death counts are based on death records and is received  by the NCHS receives the death records from the state vital registration offices through the Vital Statistics Cooperative Program (VSCP). The provisional counts are reported by the county were the decedent resided, not necessarily where they died.

County Well-Being Rankings:
The County Well-Being Rankings are for the years 2014-2023 and is collected County Health Rankings & Roadmaps CHR&R) is a program of the University of Wisconsin Population Health Institute. The health rankings include deaths which include overdose deaths, years of potential life lost, mental health statistics, income, and other health public heatlh rankings.

Overdose by Drug:
The Overdose by drug dataset contains information on non-fatal overdose data by state for 2018-2023. Provided by the CDC, DOSE (Drug Overdose Surveillance and Epidemiology) collects health department reports in each state.

Other variables of interest include opioid and stimulant overdoses and the significance of the change from previous months in those overdoses.

Violent Crime Data
From the FBI crime data reporter database, the number of violent crimes in Oregon.

How are they related? 
The Well-Being Ranks, Overdose Mortality by County contain county fips codes, county names and year in relation. The Overdose by Drug has a relationship of total non-fatal drug overdoses in each state.

Data Processing:

Findings:

Limitations:

Results:
```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```

```{r setup}

```