### 2023-11-30 Code/Draft1Rmd.Rmd
Update Draft1Rmd.Rmd

---
title: "Exploratory Analysis"
author: "James Santarpio and Mardi Elings"
date: "2023-11-30"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(ggplot2)
library(Synth)
source("Functions.R")

url <- c("https://raw.githubusercontent.com/mardimardimardi/ECNS_460_TopicData/main/Data/Cleaned%20Data/CleanedStateHealthRankings/MergedAllCountyRanks.csv")
# Read csv
data <- read.csv(url)

# Remove 2023- 2023 is a duplicate of 2022 for much of the data
health_ranking <- data %>%
  filter(year <= 2022) %>%
  filter(! (county %in% c("year_state_total")))

```
# Data Analytics Term Project
## Project Desription

### Topic:
The effect of drug decriminalization in Oregon on overdose deaths and well-being compared to California and Washington.

### Why it Matters 
The study of drug decriminalization and overdose rates in Oregon has potential for significant policy and societal relevance. There are many debates on the benefits and drawbacks that drug use and decriminalization has on potential for policy implications involving economics, public health, criminal justice, and community well-being.

### Research Question and Motivation
What is the effect of drug decriminalization on overdose deaths and well-being? A policy change like Oregon's can have significant consequences on populations who are most at risk of overdose deaths, economic impacts on health care costs, and overall well-being. Investigating the effects with county level data, we used a  difference-in-difference model with fixed effects to assess the impact of the policy change in Oregon. By comparing the outcomes in Oregon before and after the intervention to a control group of Washington and California counties as control groups.

### Description of Data

#### Overdose Mortality by County in the US:
The Overdose Mortality by County in the US is from the CDC Center for Health Statistics for the years 2020-2023. These provisional county-Level drug Overdose death counts are based on death records and is received by the NCHS receives the death records from the state vital registration offices through the Vital Statistics Cooperative Program (VSCP). The provisional counts are reported by the county were the decedent resided, not necessarily where they died.

#### County Well-Being Rankings:
The County Well-Being Rankings are for the years 2014-2023 and is collected by County Health Rankings & Roadmaps CHR&R, a program of the University of Wisconsin Population Health Institute. The health rankings include deaths which include overdose deaths, years of potential life lost, mental health statistics, income, and other health public health rankings.

#### Overdose by Drug:
The Overdose by drug dataset contains information on non-fatal overdose data by state for 2018-2023. Provided by the CDC, DOSE (Drug Overdose Surveillance and Epidemiology) collects health department reports in each state.

Other variables of interest include opioid and stimulant overdoses and the significance of the change from previous months in those overdoses.
Violent Crime Data From the FBI crime data reporter database, the number of violent crimes in Oregon.

#### How are they related? 
The Well-Being Ranks, Overdose Mortality by County contain county fips codes, county names and year in relation. The Overdose by Drug has a relationship of total non-fatal drug overdoses in each state.

## Data Processing 
For the county health rankings data, we originally had years 2014-2023, but after further inspection we found the way overdose deaths were collected for 2014-2015 was significantly different than for the years 2016-2023. We opted to drop those years and due to those data sets not including as many well-being measures.

The data collected was generally in clean condition. Some column name changes were done, and general data manipulation was completed. For example, health rankings came from yearly filet, which had to be merged together. There was generally little missing data. Violent crime and overdose by drugs contained zero missing values. However, a key data set used, health rankings, contained 13 percent N/A values. These appeared generally randomly scatter, and we will assume these are missing at random and will be ignored.

### Findings and Visualizations

Initial findings include general decrease in well being, after 2020 which is likely due to Covid-19. Well-being variables include, mental health days, fair or poor health days, excessive drinking, and income. Additionally, drug overdose deaths have been increasing, especially since 2020. However, it appears that Oregon's increase may be slower relative to Washington and Oregon after decriminalizing drugs.

We compared the parallel trends of overdose deaths per-capita in all states. Pre-treatment shows (2016-2020) shows

``` {r}
# Show deaths per-capita for each state
data2 <- health_ranking |>
  mutate(od_deaths_percapita = combined_overdose_deaths / population)

# All states and overdose death per-capita for years 2016, 2020, 2023
data2 |>
  filter(year %in% c(2016, 2018, 2020, 2023) & !is.na(od_deaths_percapita)) |>
  ggplot(aes(od_deaths_percapita, y = after_stat(count), fill = state)) +
  scale_x_continuous(
    trans = "log10",
    labels = scales::label_number(scale = 1e6, suffix = "M")  # Adjust scale based on your data
  ) +
  geom_density(alpha = 0.3, position = "stack") +
  facet_grid(year ~ .) +
  labs(
    title = "Density Plot of Overdose Deaths Per Capita by State",
    x = "Overdose Deaths Per Capita",
    y = "Count of Overdose Deaths",
    fill = "State"
  )

```

``` {r}
# Show deaths per-capita for each state
data2 <- health_ranking |>
  mutate(od_deaths_percapita = combined_overdose_deaths / population)

# All states and overdose death per-capita for years 2016, 2020, 2023
data2 |>
  filter(year %in% c(2016, 2018, 2020, 2023) & !is.na(od_deaths_percapita)) |>
  ggplot(aes(od_deaths_percapita, y = after_stat(count), fill = state)) +
  scale_x_continuous(
    trans = "log10",
    labels = scales::label_number(scale = 1e6, suffix = "M")  # Adjust scale based on your data
  ) +
  geom_density(alpha = 0.3, position = "stack") +
  facet_grid(year ~ .) +
  labs(
    title = "Density Plot of Overdose Deaths Per Capita by State",
    x = "Overdose Deaths Per Capita",
    y = "Count of Overdose Deaths",
    fill = "State"
  )

```

```{r}
#Create individual data sets for Oregon, Washington, and California
or_health_ranking <- health_ranking[health_ranking$state == "Oregon",]
ca_health_ranking <- health_ranking[health_ranking$state == "California",]
wa_health_ranking <- health_ranking[health_ranking$state == "Washington",]

#Now create yearly averages for the states from the healthrank_to_yearly function from Functions.R:
or_yearly <- healthrank_to_yearly(or_health_ranking)
ca_yearly <- healthrank_to_yearly(ca_health_ranking)
wa_yearly <- healthrank_to_yearly(wa_health_ranking)

ggplot() +
  geom_line(data = or_yearly, aes(x = year, y = log(combined_overdose_deaths), color = "Oregon")) +
  geom_line(data = ca_yearly, aes(x = year, y = log(combined_overdose_deaths), color = "California")) +
  geom_line(data = wa_yearly, aes(x = year, y = log(combined_overdose_deaths), color = "Washington")) +
  geom_point(data = or_yearly, aes(x = year, y = log(combined_overdose_deaths), color = "Oregon"), size = 3) +
  geom_point(data = ca_yearly, aes(x = year, y = log(combined_overdose_deaths), color = "California"), size = 3) +
  geom_point(data = wa_yearly, aes(x = year, y = log(combined_overdose_deaths), color = "Washington"), size = 3) +
  geom_vline(xintercept = as.numeric("2021"), linetype = "dashed", color = "black", size = 1) +
  scale_color_manual(name = "State",
                     values = c("Oregon" = "blue", "California" = "red", "Washington" = "green")) +
  labs(y = "Log Combined Overdose Deaths") +
  theme_bw()
```
```

Comparing the rate of drug mortality by county depicts increases in many counties in all three states we are examining.

```{r Cleveland dot plots, fig.height=5, fig.width=8}

# Create Cleveland dot plots for each state to show changes in county before and after the policy change 
# Arrange the data mortality in desc order and subset into states to see the changes in county by year
health_ranking_OR <- health_ranking |>
  arrange(desc(combined_overdose_mortality_rate)) |>
  filter(state == "Oregon")

health_ranking_WA <- health_ranking |>
  arrange(desc(combined_overdose_mortality_rate)) |>
  filter(state == "Washington")

health_ranking_CA <- health_ranking |>
  arrange(desc(combined_overdose_mortality_rate)) |>
  filter(state == "California")

# Oregon
health_ranking_OR %>%
  filter(year %in% c(2016, 2020, 2023) & !is.na(combined_overdose_mortality_rate)) %>%
  mutate(county = reorder(county, combined_overdose_mortality_rate)) %>%
  ggplot(aes(combined_overdose_mortality_rate, county)) +
  geom_line(aes(group = county)) +
  geom_point(size = 2, aes(color = factor(year))) +
  labs(title = "Overdose Mortality Rate in Oregon",
       x = "Overdose Mortality Rate",  # Add x-axis label
       y = "County") +
  theme_minimal()

# Washington
health_ranking_WA %>%
  filter(year %in% c(2016, 2020, 2023) & !is.na(combined_overdose_mortality_rate)) %>%
  mutate(county = reorder(county, combined_overdose_mortality_rate)) %>%
  ggplot(aes(combined_overdose_mortality_rate, county)) +
  geom_line(aes(group = county)) +
  geom_point(size = 2, aes(color = factor(year))) +
  labs(title = "Overdose Mortality Rate in Washington",
       x = "Overdose Mortality Rate",  # Add x-axis label
       y = "County") +
  theme_minimal()
```

```{r CA Cleveland dot plots, fig.height=7, fig.width=8}
# California 
# Save the ggplot to a variable

health_ranking_CA %>%
  filter(year %in% c(2016, 2020, 2023) & !is.na(combined_overdose_mortality_rate)) %>%
  mutate(county = reorder(county, combined_overdose_mortality_rate)) %>%
  ggplot(aes(combined_overdose_mortality_rate, county)) +
  geom_line(aes(group = county)) +
  geom_point(size = 2, aes(color = factor(year))) +
  labs(title = "Overdose Mortality Rate in California",
       x = "Overdose Mortality Rate",  # Add x-axis label
       y = "County") +
  theme_minimal()

```

## Methods

## Difference-in-Difference 
We used a difference-in-difference regression with fixed effects to investigate the impact of drug decriminalization starting in 2021 in Oregon counties compared to other states. The treatment effect shows a 80.39 decrease in overdose deaths compared to the control group and is statistically significant (p < 0.001). The fixed effects for years 2019 onwards are statistically signifcant, which indicate a time trend.

```{r results="hide" }
# Create a DiD variable
did_data <- health_ranking |>
  mutate(treatment = ifelse(state == "Oregon" & year >= 2021, 1, 0))

# Fit the DiD model using year and county fixed effects and population as a covariate 
did_model_county <- lm(combined_overdose_deaths ~ treatment + factor(county) + factor(year) + treatment:factor(year) + population, data = did_data)
summary(did_model_county)

```

### Synthetic Control 
A synthetic control analysis was conducted using Washington and Oregon as donor states. This analysis has significant limitations due to the lack of donor states and potential lack of credible covariates.

```{r results="hide"}
#Synthetic Control

#Create indivdual data sets for Oregon, Washington, and California
or_health_ranking <- health_ranking[health_ranking$state == "Oregon",]
ca_health_ranking <- health_ranking[health_ranking$state == "California",]
wa_health_ranking <- health_ranking[health_ranking$state == "Washington",]

healthrank_to_yearly <- function(df){
  df <- df[!df$county == "year_state_total",]
  df <- df %>%
    group_by(year) %>%
    summarise(across(where(is.numeric), ~ mean(., na.rm = TRUE))) %>%
    select(-c("X", "fips"))
  return(df)
}

#Now create yearly averages for the states from the healthrank_to_yearly function from Functions.R:
or_yearly <- healthrank_to_yearly(or_health_ranking)
ca_yearly <- healthrank_to_yearly(ca_health_ranking)
wa_yearly <- healthrank_to_yearly(wa_health_ranking)

#The three states follow trends mostly closely (Relative to each scale)- Good fit for synthetic control

#Create state number variables to use for synthetic control
health_ranking$state_num <- ifelse(health_ranking$state == "Oregon", 1, 2)

health_ranking$state_num <- ifelse(health_ranking$state == "Oregon", 1,
                                   ifelse(health_ranking$state == "California", 2, 3))

#Lets create long data set for each year- for the synth package
#Add the state numbers and states to the yearly data sets
or_yearly$state_num <- 1
or_yearly$X.1 <- "Oregon"
ca_yearly$state_num <- 2
ca_yearly$X.1 <- "California"
wa_yearly$state_num <- 3
wa_yearly$X.1 <- "Washington"

#Join the three yearly data sets by row
health_ranking_yearly <- rbind(rbind(or_yearly, ca_yearly), wa_yearly)
health_ranking_yearly <- as.data.frame(health_ranking_yearly) #for synth package purposes

#Run a synthetic control with the predictor variable being combined overdose deaths. The two states Washington and 
#California will serve as the donor pool states. Variables used will be smokers, percent fair or poor health, excessive drinking, and unemployed.
#The training data will be from 2016-2020, with 2021 and 2022 being the treatment outcomes section.
dataprep.out <-
  dataprep(health_ranking_yearly,
           predictors = c("smokers", "percent_fair_or_poor_health", "excessive_drinking"),
           dependent = "combined_overdose_deaths",
           unit.variable = "state_num",
           time.variable = "year",
           unit.names.variable = "X.1",
           treatment.identifier = 1,
           controls.identifier = c(2,3),
           time.predictors.prior = c(2016:2020),
           time.optimize.ssr = c(2016:2020),
           time.plot = 2016:2022,
  )
synth.out <- synth(dataprep.out)
print(synth.tables <- synth.tab(
  dataprep.res = dataprep.out,
  synth.res = synth.out)
)

#Create a plot of pre and post treatment
path.plot(synth.res = synth.out,
          dataprep.res = dataprep.out,
          Ylab = c("Y"),
          Xlab = c("Year"),
          Legend = c("Oregon", "Synthetic Oregon"),
          Legend.position = c("topleft")
)

```

### Results

Due to the limitations surrounding our synthetic control analysis, our synthetic Oregon was higher than the actual Oregon pre treatment which should not have happened. This may be due to a lack of accurate covariates used. Our analysis shows synthetic Oregon increasing more than actual Oregon, however limitations have to be noted surrounding this conclusion.

Diff in Diff:

Decriminalizing drugs in Oregon has slowed the increase in drug overdose deaths.

### Limitations 
Oregon drug decriminilization has recently went into effect. The data for overdose deaths after the law has been enacted is limited with only 2021 and 2022. 
It is a challenge to find a single cause for an impact on drug decriminalization and its relationship to drug overdose deaths. Since multiple factors can contribute to drug-related outcomes. There has been some evidence linked to the Covid-19 pandemic contributing to drug use, well-being, and mental health. There is limited generalizability to this study as findings from synthetic control analysis are specific to the context of Oregon drug policies and may not be generalized to other regions or policies. There is also limited data post treatment period (2 years). Additionally, the synthetic Oregon may not be a good fit.